### 2019-03-08 paper.tex
Progress

\documentclass{article}
\begin{document}
\title{A pub-sub mechanism for Cardano and Plutus}
\author{Javier Casas Velasco}
\date{March 2019}
\maketitle

\begin{abstract}
  On this paper we demonstrate how we can use DataScripts to communicate different programs using the Cardano Blockchain and the Plutus Smart Contract programming language. We construct a mechanism that works in a way akin to a pub-sub communication system.
\end{abstract}

\tableofcontents

\section{Introduction}
Blockchains provide a new way to intercommunicate different actors in the world. The blockchains provide a solution for the old problem of getting to cooperate different actors that may not want to cooperate. In this sense, the main use of the Blockchain has been so far as a ledger that is mostly handled manually. When two actors want to communicate, they use mostly manual approaches (such as chats, videoconferences, or plain old phone calls) before actually using the Blockchain, in most cases as a ledger for paying.

In order to be able to automate some of this manual communication, an automatic way of communicating these actors shall be developed, so that the actors can communicate without resorting to manual ways.

\subsection{The Publisher-Subscriber (pub-sub) pattern}

One of the main building blocks of distributed process communication is the Publisher-Subscriber (also known as pub-sub) pattern. It allows one or many broadcaster processes to send messages to one or many receiver processes. Many other patterns of distributed systems can be constructed from it.
For example, client-server communication between two processes can be achieved by having two pub-sub patterns, one for the client to send the queries to the server, the other for the server to send the responses back to the client.

The main roles in the pub-sub pattern are:
\paragraph{Publisher}
The publisher sends data messages to the channel. There may be many publishers broadcasting messages at the same time.

\paragraph{Subscriber}
The subscriber receives messages from the channel, and acts on them. There may be many subscribers receiving messages from the channel, and each subscriber receives all the messages.

\paragraph{Channel}
A channel is an abstraction that allows publishers to send messages to all the subscribers.

\paragraph{Message}
A message is an arbitrary piece of data that can be encoded and sent in a channel to be received and decoded by the subscribers listening on the channel.

\paragraph{Broadcasting}
The pub-sub pattern uses broadcasting as the single way of communication. A copy of each message sent to the channel is delivered to each subscriber.

The main difference between pub-sub and other many-to-many communication patterns is that pub-sub delivers all the messages to all the subscribers, whereas other patterns may decide to route the messages to a restricted group of receivers based on an arbitrary policy.

\subsection{The Cardano blockchain}
Cardano is a blockchain based on the Outboros Proof of Stake {QUOTE HERE}. It provides a model for universal transactions called UTxO. The Outboros Proof of Stake protocol provides protection against many of the usual attacks against blockchains, while being a Proof of Stake protocol, and thus requiring very little computational power in comparison with Proof of Work blockchains. On top of the usual cryptocurrency operations, Cardano seeks to provide other services, including Smart Contracts.

\subsection{Plutus Smart Contracts in the Cardano blockchain}
A Smart Contract is a transaction in a blockchain governed by an arbitrary program. This is different from the usual transactions, that are usually governed solely by one or several cryptographic signature.

The Smart Contracts for the Cardano blockchain will be written in a language called Plutus, and based on Haskell. At the time of writing this paper, Plutus is in heavy evolution, and lacks yet a testing network. The only way to execute Plutus smart contracts is through the Plutus Emulator, or through the Plutus Playground, which uses the emulator under the hood.

The way Plutus Smart Contracts are implemented is quite simple, but requires some understanding:

\begin{itemize}
\item A stakeholder pays an amount to the address of a smart contract (which is the hash of the smart contract). In this payment, the stakeholder provides a DataScript: a piece of data in the Plutus format that can be used to configure the smart contract.
\item A second stakeholder wants to spend the amount the first stakeholder paid to the smart contract address. To do so, he has to provide a smart contract which hash coincides with the first address, and a second piece of data called the RedeemerScript, and, when executing the smart contract code with the DataScript, the RedeemerScript and the transaction as parameters, it has to terminate without any errors. Then the transaction is authorised.
\end{itemize}

This implementation provides several interesting approaches that we are going to use to construct our pub-sub mechanism.

\section{Publishing and receiving messages}

In order to create a pub-sub mechanism, we need a way to write messages to the Cardano blockchain, and a mechanism to recover messages from the blockchain. The blockchain is public, and we will use this property to derive the broadcasting property of pub-sub from it.

\subsection{Publishing messages to the blockchain}
First, we need to publish messages to the blockchain. For this, we will use smart contracts, but not in a complete way. We will encode our messages in DataScripts, and will publish them to the blockchain by paying to the address of a Smart Contract with the DataScript as attached data.

For this, we can use the standard mechanisms that Plutus provides:

\begin{description}
\item [Ledger.Types.lifted] converts a bunch of data to Plutus format.
\item [Wallet.API.payToScript] pays an amount of Ada to the address of a Smart contract with the DataScript provided.
\end{description}

A usual implementation of this can be:

\begin{verbatim}
CODE GOES HERE
\end{verbatim}

\subsection{Receiving messages from the blockchain}
Receiving messages is also supported by the blockchain, but not in the way the Cardano team envisioned.

The Wallet API supports a wallet listening for activity in arbitrary addresses. This is intended for the wallet to be notified when it receives a transaction, or when activity occurs at a smart contract of interest. This mechanism is based on two Wallet API endpoints:

\begin{description}
\item [Wallet.API.startWatching] starts watching the specified address, and adds any transactions that happen on it to the dictionary of watched addresses.
\item [Wallet.API.watchedAddresses] returns the dictionary of watched addresses, that includes all the associated transactions for all the watched addresses, including DataScripts for associated payToScript transactions.
\end{description}

This means we can recover data from the blockchain, but in the format of Plutus smart contracts. The messages will be accumulated on the AddressMap in arbitrary order, and this is something we have to keep in mind when constructing our systems.

\section{Converting arbitrary data from and to DataScripts}
The blockchain will accept and return data in the Plutus format, but we want to process it as Haskell data, so our standard Haskell applications can just use it.

\subsection{Converting arbitrary data into DataScripts}
Converting arbitrary data to a DataScript is well supported in the system. As stated previously,

\begin{description}
\item [Ledger.Types.lifted] converts a bunch of data to Plutus format.
\end{description}

The only requirement is that the data has a Language.PlutusTx.Lift.Class.Lift instance, which can be manufactured automatically by using the Language.PlutusTx.makeLift Template Haskell incantation.

\begin{verbatim}
module MyModule where

import qualified Language.PlutusTx            as PlutusTx
import           Ledger                       (Slot(..), PubKey)
import           Ledger.Ada.TH                (Ada)

-- Declare our datatype
data MyDatatype = MyDatatype
    { foo :: Slot
    , bar :: Int
    , baz :: Ada
    }

-- Create a Language.PlutusTx.Lift.Class.Lift instance for it
PlutusTx.makeLift ''MyDatatype
\end{verbatim}

\subsection{Recovering arbitrary data from DataScripts}
Lifting data to the Plutus format is as easy as using the Ledger.Types.lifted function. But the way back is not that easy.

\section{Conclusions}

    Hello, world!

    Instead of WYSIWYG editors, typesetting systems like \TeX{} or \LaTeX{} \cite{lamport94} can be used.

\begin{thebibliography}{9}

\bibitem{lamport94}
  Leslie Lamport,
    \textit{\LaTeX: a document preparation system},
    Addison Wesley, Massachusetts,
    2nd edition,
    1994.

\end{thebibliography}

\end{document}